---
title: "Extracting trail and road data from OpenStreetMap"
author: "Forest P Hayes"
date: "December, 16, 2020"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, echo = F}
library(magrittr)
```

## 1. Define extent
#### Create a bounding box for the study area
```{r}
glac_bbox <- 
  # approximate bounding box for Glacier National Park
  sf::st_bbox(c(xmin = -114.47552, ymin = 48.23369, xmax = -113.24198, ymax = 49.0000))

```


## 2. Download data

#### Get trail data
Note: this includes the classifications of 'path' (multi-use), 'footway' (foot traffic only), and "steps" (i.e. stairs)
```{r}
trails <- 
  osmdata::opq(bbox = glac_bbox) %>% 
  osmdata::add_osm_feature(key = 'highway', value = c("path", "footway", "steps")) %>% 
  osmdata::osmdata_sf()

```


#### Get trailhead data
Note: this is specifically for _named_ trailheads. There are numerous parking areas that are at the start of trails that do now show up as trailheads.
```{r}
trailhead <- 
  osmdata::opq(bbox = glac_bbox) %>% 
  osmdata::add_osm_feature(key = 'highway', value = c("trailhead")) %>% 
  osmdata::osmdata_sf()

```


#### Get parking area data
This seems like it could be useful to augment trailhead data. That said, parking areas include many pull outs along roads that are deffinitly _not_ trail access locations. 


```{r}
# parking <-
#   osmdata::opq(bbox = glac_bbox) %>%
#   osmdata::add_osm_feature(key = 'amenity', value = c("parking")) %>%
#   osmdata::osmdata_sf()

```



#### Get data for all road types
Note the number of different road classifications that OpenStreetMap uses. All common road types are included below.
```{r}
all_roads <-
    osmdata::opq(bbox = glac_bbox) %>%
    osmdata::add_osm_feature(
      key = 'highway', 
      value = c("motorway", "trunk", "primary", "secondary", "tertiary", 
                "unclassified", "residential", "motorway_link", "trunk_link", 
                "primary_link", "secondary_link", "tertiary_link", 
                "living_street", "service")
    ) %>%
    osmdata::osmdata_sf()

```

## 3. Map data

```{r, out.width="100%"}
mapview::mapview(
  list(
    Roads = all_roads$osm_lines,
    Trail = trails$osm_lines,
    Trailhead = trailhead$osm_points
  ),
  color = list("saddlebrown", "coral", "darkblue"),
  col.regions = "blue"
)@map

```

